\textbf{Evaluation of machine learning methods for 16S rRNA gene data}

### 2019-05-22 submission/manuscript.tex
update to figures and paper and abstract:

Advances in sequencing technology and decreasing costs of generating 16S
rRNA gene sequences have allowed rapid exploration of human associated
microbiome and its health implications. Currently, the human microbiome
field is growing at an unprecedented rate and as a result, there is an
increasing demand for methods that identify associations between members
of the microbiome and human health. However, this is difficult as human
associated microbial communities are remarkably complex and uneven. It
is unlikely that a single species can explain a disease. Instead,
subsets of those communities, in relation to one another and to their
host, account for the differences in health outcomes. Machine learning
(ML) methods are effective at recognizing and highlighting patterns in
complex microbial datasets. Therefore, researchers have started to
explore the utility of ML models that use microbiota associated
biomarkers to predict human health and to understand the microbial
ecology of diseases such as liver cirrhosis, colorectal cancer,
inflammatory bowel diseases (IBD), obesity, type 2 diabetes and others
(1--11). However, currently the field's use of ML lacks clarity and
consistency on which methods are used and how these methods are
implemented {[}(12); . Most notably, there are misleading practices of
using ``leaky'' ML pipelines where models are tested on training data,
reporting few or only the best outcomes of different randomizations of
cross-validation and showing a disregard for large differences between
cross-validation and testing performances as well as large confidence
intervals of testing performances (4, 13--18). Moreover, there is a lack
of discussion on why a particular ML model is utilized. Recently, there
is a trend towards using more complex ML models such as random forest,
extreme gradient boosting and neural networks without a discussion on if
and how much model interpretibility is necessary for the study (11,
19--21). The lack of transparency on modeling methodology and model
selection negatively impact model reproducibility and reliability. We
need to strive toward better machine learning practices by (1)
implementing consistent and reliable machine learning pipelines; (2)
selecting ML models that reflect the goal of the study as it will inform
our expectations of model accuracy, complexity, interpretibility and
computational efficiency.

@@

We used a cohort of 490 patients with 261 cases of SRN. For each
patient, we had 6920 features (fecal bacterial abundances) and a
two-class label that defines their colonic health (SRN or normal). All
the cases were independently labeled through colonoscopies. We
established modeling pipelines for a binary prediction task with
L2-regularized logistic regression, L1 and L2 support vector machines
(SVM) with linear and radial basis function kernels, a decision tree,
random forest and XGBoost to emphasize the differences in model
accuracy, complexity, interpretibility and computational efficiency due
to model selection. We randomly split the data into training/validation
and test sets so that the training/validation set consisted of 80\% of
the full dataset while the test set was composed of the remaining data
{[}Figure 1{]}. Since the cases are not uniformly represented in the
data, the initial data-split was stratified to maintain the overall
label distribution on both the training/validation and test sets.
Training/validation set consisted of 393 patients (209 SRN), while the
test set was composed of 97 patients (52 SRN). The training/validation
data was used for training purposes and validation of hyperparameter
selection, and the test set was used for evaluation purposes. Validation
of hyperparameter selection was performed using 100 randomizations of
five-fold cross-validation on the training/validation set {[}Figure
1{]}. Similar to the initial data-split, five-fold cross-validation was
also stratified to maintain the overall label distribution on the
training and validation sets. We validated the cross-validation
performances of each hyperparameter setting over the 100 randomizations
and selected the best performing hyper-parameter setting to train the
full training/validation dataset {[}Figures S1 and S2{]}. We then used
the held-out test set to evaluate the prediction performance of each ML
model. The data-split, hyperparameter selection, training and testing
steps were repeated 100 times to get a reliable and robust reading of
model prediction performance {[}Figure 1{]}.

\textbf{Random forest has the highest prediction performance when
applied to the held-out test data compared the other six ML models .}

@@

\textbf{Predictive performances of the seven ML models during
cross-validation and when applied to the held-out test data do not show
large differences.}

We evaluated the generalization performance of each ML model by
comparing their median cross-validation AUROC and median testing AUROC.
The largest difference between the two was 0.021 in L1 SVM with linear
kernel followed by SVM with radial basis function kernel and decision
tree with a difference of 0.007 and 0.006, respectively.

\textbf{The complexity and interpretibility of each ML model.}

The ML models we built using L2-regularized logistic regression, L1 and
L2 support vector machines (SVM) with linear and radial basis function
kernels, a decision tree, random forest and XGBoost increase in
complexity as they decrase in interpretibility. We interpreted L1 and L2
SVM with linear kernel using the feature weights and L2 logistic
regression using regression coefficients of the trained models. We
calculated the mean weights and coefficients of all the features over
the 100 data-splits. In the three linear models, OTUs that had the
largest mean weights and drove the detection of SRNs belong to family
\emph{Lachnospiraceae}, and \emph{Ruminococcaceae} (OTU01239, OTU00659,
OTU00742, OTU00012, OTU00050, OTU00015, OTU00768, OTU00822, OTU00609,
OTU01212, OTU00629) and genera \emph{Gamella} (OTU00426) {[}Figure 3{]}.
We explained the feature importances in non-linear models using
permutation importance on the held-out test data. For the tree-based
models, permuting \emph{Peptostreptococcus} (OTU00367) abundances
randomly, dropped the predictive performances the most. Decision tree,
random forest and XGBoost models' predictive performance dropped from
0.6 base testing AUROC median to 0.52, from 0.69 to 0.68 and from 0.68
to 0.65, respectively {[}Figure 4{]}. The negative predictive impact of
\emph{Peptostreptococcus} in the decision tree model was followed by a
\emph{Lachnospiraceae} species (OTU00058) (0.6 base testing AUROC median
to 0.58) {[}Figure 4B{]}. Other OTUs had none to minimal effect on the
predictive performance.

@@

Our results show that Suprisingly, l2 logistic.. computational
efficiency and interp of logistic vs random forest decision tree and l1
why so low?

@@

What happens with imbalanced data What happens with smaller datasets
Does the code work on any microbiome data? Why did we use only one
dataset? Becuase we are comparing several methods on 1 dataset. What
will the future data look like?

@@

The data used for this analysis are stool bacterial abundances and
clinical information of the patients recruited by Great Lakes-New
England Early Detection Research Network study. These data were obtained
from Sze et al (22). The stool samples were provided by recruited adult
participants who were undergoing scheduled screening or surveillance
colonoscopy. Colonoscopies were performed and fecal samples were
collected from participants in four locations: Toronto (ON, Canada),
Boston (MA, USA), Houston (TX, USA), and Ann Arbor (MI, USA). Patients'
colonic health was labeled by colonoscopy with adequate preparation and
tissue histopathology of all resected lesions. Patients with an adenoma
greater than 1 cm, more than three adenomas of any size, or an adenoma
with villous histology were classified as advanced adenoma. Study had
172 patients with normal colonoscopies, 198 with adenomas and 120 with
carcinomas. Of the 198 adenomas, 109 were identified as advanced
adenomas. Stool provided by the patients was used for 16S rRNA gene
sequencing to measure bacterial population abundances. The bacterial
abundance data was generated by Sze et al, by processing 16S rRNA
sequences in Mothur (v1.39.3) using the default quality filtering
methods, identifying and removing chimeric sequences using VSEARCH and
assigning to OTUs at 97\% similarity using the OptiClust algorithm
(23--25).

@@

\includegraphics{Figure_1} \textbf{Figure 1. Machine learning pipeline
showing predictive model training and evaluation flowchart. } We split
the data 80\%/20\% stratified to maintain the overall label
distribution, performed five-fold cross-validation on the training data
to select the best hyperparameter setting and then using these
hyperparameters to train all of the training data. The model was
evaluated on a held-out set of data (not used in selecting the model).
Abbreviations: AUROC, area under the receiver operating characteristic
curve \newpage
\includegraphics{Figure_2.png} \textbf{Figure 2. Generalization and
classification performance of ML models using AUROC values of all cross
validation and testing performances. } The median AUROC for diagnosing
individuals with SRN using bacterial abundances was higher than chance
(depicted by horizontal line at 0.50) for all the ML models.
Discriminative performance of random forest model was higher than other
ML models. The boxplot shows quartiles at the box ends and the
statistical median as the horizontal line in the box. The whiskers show
the farthest points that are not outliers. Outliers are data points that
are not within 3/2 times the interquartile ranges. Abbreviations: SRN,
screen-relevant neoplasias; AUROC, area under the receiver operating
characteristic curve; SVM, support vector machine; XGBoost, extreme
gradient boosting \newpage
\includegraphics{Figure_3.png} \textbf{Figure 3. Interpretation of the
linear ML models.} (A) L2 logistic regression coefficients (B) L1 SVM
with linear kernel feature weights (C) L2 SVM with linear kernel feature
weights. The means weights and coefficients of the most important five
OTUs for each model are shown here with the standard deviation over 100
data-splits. Similar OTUs had the largest impact on the predictive
performance of L2 logistic regression and L2 SVM with linear kernel.
Abbreviations: SVM, support vector machine; OTU, Operational Taxonomic
Unit. \newpage
\includegraphics{Figure_4.png} \textbf{Figure 4. Explanation of the
non-linear ML models.} (A) SVM with radial basis kernel (B) decision
tree (C) random forest (D) XGboost feaure importances were explanied
using permutation importance using held-out test set. The gray rectangle
and the dashed line show the IQR range and median of the base testing
AUROC without any permutation performed. For all the tree-based models,
a \emph{Peptostreptococcus} species (OTU00367) had the largest impact on
predictive performance of the model. Abbreviations: SVM, support vector
machine; OTU, Operational Taxonomic Unit; RBF, radial basis kernel; OTU,
Operational Taxonomic Unit. \newpage
\includegraphics{Figure_5.png} \textbf{Figure 5. Computational
efficiency of seven ML models.} The walltimes for training and testing
of each data-split showed the differences in computational efficieny of
the seven models. The median walltime in hours was the highest for
XGBoost and shortest for L2 logistic regression. The boxplot shows
quartiles at the box ends and the statistical median as the horizontal
line in the box. The whiskers show the farthest points that are not
outliers. Outliers are data points that are not within 3/2 times the
interquartile ranges. Abbreviations: AUROC, area under the receiver
operating characteristic curve; SVM, support vector machine; XGBoost,
extreme gradient boosting.\\
\newpage

@@

\hypertarget{ref-galkin_human_2018}{}
19. \textbf{Galkin F}, \textbf{Aliper A}, \textbf{Putin E},
\textbf{Kuznetsov I}, \textbf{Gladyshev VN}, \textbf{Zhavoronkov A}.
2018. Human microbiome aging clocks based on deep learning and tandem of
permutation feature importance and accumulated local effects. bioRxiv.
doi:\href{https://doi.org/10.1101/507780}{10.1101/507780}.

\hypertarget{ref-reiman_using_2017}{}
20. \textbf{Reiman D}, \textbf{Metwally A}, \textbf{Dai Y}. 2017. Using
convolutional neural networks to explore the microbiome, pp. 4269--4272.
\emph{In} 2017 39th annual international conference of the IEEE
engineering in medicine and biology society (EMBC).

\hypertarget{ref-fioravanti_phylogenetic_2017}{}
21. \textbf{Fioravanti D}, \textbf{Giarratano Y}, \textbf{Maggio V},
\textbf{Agostinelli C}, \textbf{Chierici M}, \textbf{Jurman G},
\textbf{Furlanello C}. 2017. Phylogenetic convolutional neural networks
in metagenomics. arXiv:170902268 {[}cs, q-bio{]}.

\hypertarget{ref-sze_leveraging_2018}{}
22. \textbf{Sze MA}, \textbf{Schloss PD}. 2018. Leveraging existing 16S
rRNA gene surveys to identify reproducible biomarkers in individuals
with colorectal tumors. mBio \textbf{9}:e00630--18.
doi:\href{https://doi.org/10.1128/mBio.00630-18}{10.1128/mBio.00630-18}.

\hypertarget{ref-schloss_introducing_2009}{}
23. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Van Horn DJ}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-Source, Platform-Independent,
Community-Supported Software for Describing and Comparing Microbial
Communities. ApplEnvironMicrobiol \textbf{75}:7537--7541.

\hypertarget{ref-westcott_opticlust_2017}{}
24. \textbf{Westcott SL}, \textbf{Schloss PD}. 2017. OptiClust, an
Improved Method for Assigning Amplicon-Based Sequence Data to
Operational Taxonomic Units. mSphere \textbf{2}.
doi:\href{https://doi.org/10.1128/mSphereDirect.00073-17}{10.1128/mSphereDirect.00073-17}.

\hypertarget{ref-rognes_vsearch_2016}{}
25. \textbf{Rognes T}, \textbf{Flouri T}, \textbf{Nichols B},
\textbf{Quince C}, \textbf{Mahé F}. 2016. VSEARCH: A versatile open
source tool for metagenomics. PeerJ \textbf{4}:e2584.
doi:\href{https://doi.org/10.7717/peerj.2584}{10.7717/peerj.2584}.

>>>>

Advances in sequencing technology and decreasing costs of generating 16S
rRNA gene sequences have allowed rapid exploration of human associated
microbiome and its health implications. Currently, the human microbiome
field is growing at an unprecedented rate and as a result, there is an
increasing demand for methods that identify associations between members
of the microbiome and human health. However, this is difficult as human
associated microbial communities are remarkably complex and uneven. It
is unlikely that a single species can explain a disease. Instead,
subsets of those communities, in relation to one another and to their
host, account for the differences in health outcomes. Machine learning
(ML) methods are effective at recognizing and highlighting patterns in
complex microbial datasets. Therefore, researchers have started to
explore the utility of ML models that use microbiota associated
biomarkers to predict human health and to understand the microbial
ecology of diseases such as liver cirrhosis, colorectal cancer,
inflammatory bowel diseases (IBD), obesity, type 2 diabetes and others
(1--11). However, currently the field's use of ML lacks clarity and
consistency on which methods are used and how these methods are
implemented {[}(12); . More notably, flawed ML practices are prevalent
such as using ML pipelines where there is no seperate held-out test
dataset to evaulate model performance, reporting few or only the best
outcomes of different randomizations of cross-validation and showing a
disregard for large differences between cross-validation and testing
performances as well as large confidence intervals of testing
performances (4, 13--19). Moreover, there is a lack of discussion on why
a particular ML model is utilized. Recently, there is a trend towards
using more complex ML models such as random forest, extreme gradient
boosting and neural networks without a discussion on if and how much
model interpretibility is necessary for the study (11, 20--22). The lack
of transparency on modeling methodology and model selection negatively
impact model reproducibility and reliability. We need to strive toward
better machine learning practices by (1) implementing consistent and
reliable machine learning pipelines; (2) selecting ML models that
reflect the goal of the study as it will inform our expectations of
model accuracy, complexity, interpretibility and computational
efficiency.

@@

We used a cohort of 490 patients with 261 cases of SRN. For each
patient, we had 6920 features (fecal bacterial abundances) and a
two-class label that defines their colonic health (SRN or normal). All
the cases were independently labeled through colonoscopies. We
established modeling pipelines for a binary prediction task with
L2-regularized logistic regression, L1 and L2 support vector machines
(SVM) with linear and radial basis function kernels, a decision tree,
random forest and extreme gradient boosted decision tree (XGBoost) to
emphasize the differences in model accuracy, complexity,
interpretibility and computational efficiency due to model selection.

For regularized logistic regression and SVM with linear kernel we used
L2 regularization to keep all potentially important features. For
comparison, we also trained an L1 regularized SVM model with linear
kernel. L1-regularization on microbiome data lead to a sparser solution
(i.e., force many coefficients to zero), removing features that could be
correlated with other important features, making it more difficult
include the relationships between the microbes and possibly dropping
features that are important in detecting SRNs. Finally, to explore the
potential for non-linear relationships among features and the outcome of
interest, we trained tree based models, decision tree, random forest and
XGboost, as well as an SVM with non-linear kernel.

We established a robust and reliable ML pipeline where we train and
validate each of the seven models {[}Figure 1{]}. We randomly split the
data into training/validation and test sets so that the
training/validation set consisted of 80\% of the full dataset while the
test set was composed of the remaining data {[}Figure 1{]}. Since the
cases are not uniformly represented in the data, the initial data-split
was stratified to maintain the overall label distribution on both the
training/validation and test sets. Training/validation set consisted of
393 patients (209 SRN), while the test set was composed of 97 patients
(52 SRN). The training/validation data was used for training purposes
and validation of hyperparameter selection, and the test set was used
for evaluation purposes. Validation of hyperparameter selection was
performed using 100 randomizations of five-fold cross-validation on the
training/validation set {[}Figure 1{]}. Similar to the initial
data-split, five-fold cross-validation was also stratified to maintain
the overall label distribution on the training and validation sets. We
validated the cross-validation performances of each hyperparameter
setting over the 100 randomizations and selected the best performing
hyper-parameter setting to train the full training/validation dataset
{[}Figures S1 and S2{]}. We then used the held-out test set to evaluate
the prediction performance of each ML model. The data-split,
hyperparameter selection, training and testing steps were repeated 100
times to get a reliable and robust reading of model prediction
performance {[}Figure 1{]}.

\textbf{Model performance and generalizability.}

@@

We evaluated the generalization performance of each ML model by
comparing their median cross-validation AUROC and median testing AUROC.
This difference should be low to suggest the model is not overfitting
despite the large number of features. The largest difference between the
two was 0.021 in L1 SVM with linear kernel followed by SVM with radial
basis function kernel and decision tree with a difference of 0.007 and
0.006, respectively {[}Figure 2{]}.

\textbf{Interpretation of each ML model.}

The ML models we built using L2-regularized logistic regression, L1 and
L2 support vector machines (SVM) with linear and radial basis function
kernels, a decision tree, random forest and XGBoost increase in
complexity as they decrase in interpretibility. We interpreted L1 and L2
SVM with linear kernel using the feature weights and L2 logistic
regression using regression coefficients of the trained models. We
examined the 5 predictive (highest weight) and protective (lowest
weight) features identified by the model. We calculated the mean weights
and coefficients of these features over the 100 data-splits. In the
three linear models, OTUs that had the largest mean weights and drove
the detection of SRNs belong to family \emph{Lachnospiraceae}, and
\emph{Ruminococcaceae} (OTU01239, OTU00659, OTU00742, OTU00012,
OTU00050, OTU00015, OTU00768, OTU00822, OTU00609, OTU01212, OTU00629)
and genera \emph{Gamella} (OTU00426) {[}Figure 3{]}. We explained the
feature importances in non-linear models using permutation importance on
the held-out test data. For the tree-based models, permuting
\emph{Peptostreptococcus} (OTU00367) abundances randomly, dropped the
predictive performances the most. Decision tree, random forest and
XGBoost models' predictive performance dropped from 0.6 base testing
AUROC median to 0.52, from 0.69 to 0.68 and from 0.68 to 0.65,
respectively {[}Figure 4{]}. The negative predictive impact of
\emph{Peptostreptococcus} in the decision tree model was followed by a
\emph{Lachnospiraceae} species (OTU00058) (0.6 base testing AUROC median
to 0.58) {[}Figure 4B{]}. Other OTUs had none to minimal effect on the
predictive performance.

@@

In this study we established a robust ML pipeline to use 16S rRNA
sequence counts to predict a binary health outcome. We showed the
importance of held-out test set and reporting the results of many
iterations of cross-validation and testing. Each data-split has a weight
that comes from random splitting. Seeing that testing and
cross-validation median differences being low show that the models are
generalizable and the pipeline works. Our results also show that we can
choose different models for different reasons. Getting feature
importances as weights and regression coefficients is the fastest and
easiest interpratation. Similarly, linear models require less
computational burden, fastest to train. THinking about the
applicability, and actionability of a model in clinical setting, we
might want faster results.

Suprisingly, l2 logistic regression does really well. computational
efficiency and interp of logistic vs random forest decision tree and l1
why so low?

@@

What happens with imbalanced data What happens with smaller datasets
Does the code work on any microbiome data? Why did we use only one
dataset? Why not validate on other data instead of held-out testing set?
Becuase we are comparing several methods on 1 dataset. What will the
future data look like?

@@

The data used for this analysis are stool bacterial abundances and
clinical information of the patients recruited by Great Lakes-New
England Early Detection Research Network study. These data were obtained
from Sze et al (23). The stool samples were provided by recruited adult
participants who were undergoing scheduled screening or surveillance
colonoscopy. Colonoscopies were performed and fecal samples were
collected from participants in four locations: Toronto (ON, Canada),
Boston (MA, USA), Houston (TX, USA), and Ann Arbor (MI, USA). Patients'
colonic health was labeled by colonoscopy with adequate preparation and
tissue histopathology of all resected lesions. Patients with an adenoma
greater than 1 cm, more than three adenomas of any size, or an adenoma
with villous histology were classified as advanced adenoma. Study had
172 patients with normal colonoscopies, 198 with adenomas and 120 with
carcinomas. Of the 198 adenomas, 109 were identified as advanced
adenomas. Stool provided by the patients was used for 16S rRNA gene
sequencing to measure bacterial population abundances. The bacterial
abundance data was generated by Sze et al, by processing 16S rRNA
sequences in Mothur (v1.39.3) using the default quality filtering
methods, identifying and removing chimeric sequences using VSEARCH and
assigning to OTUs at 97\% similarity using the OptiClust algorithm
(24--26).

@@

\includegraphics{Figure_1} \textbf{Figure 1. Machine learning pipeline
showing predictive model training and evaluation flowchart. } We split
the data 80\%/20\% stratified to maintain the overall label
distribution, performed five-fold cross-validation on the training data
to select the best hyperparameter setting and then using these
hyperparameters to train all of the training data. The model was
evaluated on a held-out set of data (not used in selecting the model).
Abbreviations: AUROC, area under the receiver operating characteristic
curve \newpage
\includegraphics{Figure_2.png} \textbf{Figure 2. Generalization and
classification performance of ML models using AUROC values of all cross
validation and testing performances. } The median AUROC for diagnosing
individuals with SRN using bacterial abundances was higher than chance
(depicted by horizontal line at 0.50) for all the ML models.
Discriminative performance of random forest model was higher than other
ML models. The boxplot shows quartiles at the box ends and the
statistical median as the horizontal line in the box. The whiskers show
the farthest points that are not outliers. Outliers are data points that
are not within 3/2 times the interquartile ranges. Abbreviations: SRN,
screen-relevant neoplasias; AUROC, area under the receiver operating
characteristic curve; SVM, support vector machine; XGBoost, extreme
gradient boosting \newpage
\includegraphics{Figure_3.png} \textbf{Figure 3. Interpretation of the
linear ML models.} (A) L2 logistic regression coefficients (B) L1 SVM
with linear kernel feature weights (C) L2 SVM with linear kernel feature
weights. The means weights and coefficients of the most important five
OTUs for each model are shown here with the standard deviation over 100
data-splits. Similar OTUs had the largest impact on the predictive
performance of L2 logistic regression and L2 SVM with linear kernel.
Abbreviations: SVM, support vector machine; OTU, Operational Taxonomic
Unit. \newpage
\includegraphics{Figure_4.png} \textbf{Figure 4. Explanation of the
non-linear ML models.} (A) SVM with radial basis kernel (B) decision
tree (C) random forest (D) XGboost feaure importances were explanied
using permutation importance using held-out test set. The gray rectangle
and the dashed line show the IQR range and median of the base testing
AUROC without any permutation performed. For all the tree-based models,
a \emph{Peptostreptococcus} species (OTU00367) had the largest impact on
predictive performance of the model. Abbreviations: SVM, support vector
machine; OTU, Operational Taxonomic Unit; RBF, radial basis kernel; OTU,
Operational Taxonomic Unit. \newpage
\includegraphics{Figure_5.png} \textbf{Figure 5. Computational
efficiency of seven ML models.} The walltimes for training and testing
of each data-split showed the differences in computational efficieny of
the seven models. The median walltime in hours was the highest for
XGBoost and shortest for L2 logistic regression. The boxplot shows
quartiles at the box ends and the statistical median as the horizontal
line in the box. The whiskers show the farthest points that are not
outliers. Outliers are data points that are not within 3/2 times the
interquartile ranges. Abbreviations: AUROC, area under the receiver
operating characteristic curve; SVM, support vector machine; XGBoost,
extreme gradient boosting.\\
\newpage
\includegraphics{Figure_S1.png} \textbf{Figure S1. Hyperparameter
setting performances for linear models.} Abbreviations: \newpage
\includegraphics{Figure_S2.png} \textbf{Figure S2. Hyperparameter
setting performances for non-linear models.} Abbreviations: \newpage

@@

\hypertarget{ref-wong_quantitation_2017}{}
19. \textbf{Wong SH}, \textbf{Kwong TNY}, \textbf{Chow T-C}, \textbf{Luk
AKC}, \textbf{Dai RZW}, \textbf{Nakatsu G}, \textbf{Lam TYT},
\textbf{Zhang L}, \textbf{Wu JCY}, \textbf{Chan FKL}, \textbf{Ng SSM},
\textbf{Wong MCS}, \textbf{Ng SC}, \textbf{Wu WKK}, \textbf{Yu J},
\textbf{Sung JJY}. 2017. Quantitation of faecal fusobacterium improves
faecal immunochemical test in detecting advanced colorectal neoplasia.
Gut \textbf{66}:1441--1448.
doi:\href{https://doi.org/10.1136/gutjnl-2016-312766}{10.1136/gutjnl-2016-312766}.

\hypertarget{ref-galkin_human_2018}{}
20. \textbf{Galkin F}, \textbf{Aliper A}, \textbf{Putin E},
\textbf{Kuznetsov I}, \textbf{Gladyshev VN}, \textbf{Zhavoronkov A}.
2018. Human microbiome aging clocks based on deep learning and tandem of
permutation feature importance and accumulated local effects. bioRxiv.
doi:\href{https://doi.org/10.1101/507780}{10.1101/507780}.

\hypertarget{ref-reiman_using_2017}{}
21. \textbf{Reiman D}, \textbf{Metwally A}, \textbf{Dai Y}. 2017. Using
convolutional neural networks to explore the microbiome, pp. 4269--4272.
\emph{In} 2017 39th annual international conference of the IEEE
engineering in medicine and biology society (EMBC).

\hypertarget{ref-fioravanti_phylogenetic_2017}{}
22. \textbf{Fioravanti D}, \textbf{Giarratano Y}, \textbf{Maggio V},
\textbf{Agostinelli C}, \textbf{Chierici M}, \textbf{Jurman G},
\textbf{Furlanello C}. 2017. Phylogenetic convolutional neural networks
in metagenomics. arXiv:170902268 {[}cs, q-bio{]}.

\hypertarget{ref-sze_leveraging_2018}{}
23. \textbf{Sze MA}, \textbf{Schloss PD}. 2018. Leveraging existing 16S
rRNA gene surveys to identify reproducible biomarkers in individuals
with colorectal tumors. mBio \textbf{9}:e00630--18.
doi:\href{https://doi.org/10.1128/mBio.00630-18}{10.1128/mBio.00630-18}.

\hypertarget{ref-schloss_introducing_2009}{}
24. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Van Horn DJ}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-Source, Platform-Independent,
Community-Supported Software for Describing and Comparing Microbial
Communities. ApplEnvironMicrobiol \textbf{75}:7537--7541.

\hypertarget{ref-westcott_opticlust_2017}{}
25. \textbf{Westcott SL}, \textbf{Schloss PD}. 2017. OptiClust, an
Improved Method for Assigning Amplicon-Based Sequence Data to
Operational Taxonomic Units. mSphere \textbf{2}.
doi:\href{https://doi.org/10.1128/mSphereDirect.00073-17}{10.1128/mSphereDirect.00073-17}.

\hypertarget{ref-rognes_vsearch_2016}{}
26. \textbf{Rognes T}, \textbf{Flouri T}, \textbf{Nichols B},
\textbf{Quince C}, \textbf{Mahé F}. 2016. VSEARCH: A versatile open
source tool for metagenomics. PeerJ \textbf{4}:e2584.
doi:\href{https://doi.org/10.7717/peerj.2584}{10.7717/peerj.2584}.
